---
title: "P8105 Homework 1"
author: "Christine Lucille Kuryla"
date: "2023-09-17"
output: github_document
---

The following document is the first homework for P8105.

# Problem 1

First, load `early_january_weather` from the `penguins` dataset in the library `moderndive` as well as the tidyverse.
```{r problem1_load_and_explore, message = FALSE}

# Load moderndive library 
library(moderndive)

# Load tidyverse for dplyr to enable pull when calculating the mean temperature
library(tidyverse)

# This laods the early_january_weather dataset into a variable called early_january_weather
data("early_january_weather")

```

* The variables in this data set were: `r colnames(early_january_weather)`.
  * Some important variables included:
    * `origin`, which was `r pull(early_january_weather,origin)[1]` for every measurement.
    * Year (`year`) and month (`month`), which were the same for every observation: `r pull(early_january_weather,year)[1]` and `r pull(early_january_weather,month)[1]`, respectively.
    * Several weather measurements were included, such as
      * Temperature (`temp`), which ranged from `r min(pull(early_january_weather, temp))` to `r max(pull(early_january_weather, temp))`.
      * Dew point (`dewp`), which ranged from `r min(pull(early_january_weather, dewp))` to `r max(pull(early_january_weather, dewp))`.
      * Humidity (`humid`), which ranged from `r min(pull(early_january_weather, humid))` to `r max(pull(early_january_weather, humid))`.
      * Details about the wind, such as the direction (`wind_dir`), speed (`wind_speed`), and gusts (`wind_gust`).
      * Precipitation (`precip`), which ranged from `r min(pull(early_january_weather, precip))` to `r max(pull(early_january_weather, precip))`.
      * Pressure (`pressure`), which which had some missing values but ranged from `r min(na.omit(pull(early_january_weather, pressure)))` to `r max(na.omit(pull(early_january_weather, pressure)))`.
      * Visibility (`visib`), which had some missing values but ranged from `r min(na.omit(pull(early_january_weather, visib)))` to `r max(na.omit(pull(early_january_weather, visib)))`.
* The data set contained `r nrow(early_january_weather)` observations (rows) with `r ncol(early_january_weather)` variables (columns) describing each observation.
* The mean temperature was `r mean(pull(early_january_weather, temp))`.

```{r problem1_plot}

# Create a scatterplot of temp (y) vs time_hour (x) colored using the humid variable

ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point()

# Save the plot

ggsave("scatter_plot.pdf")

```

In the plot of the temperature vs the date/time, we can see that:

* The data measurements exist from `r min(pull(early_january_weather, time_hour))` to `r max(pull(early_january_weather, time_hour))` and the temperature ranges from `r min(pull(early_january_weather, temp))` to `r max(pull(early_january_weather, temp))`. 
* There is an overall upwards trend in the data, so it seems that as time passes from early to mid January, the temperature warms overall. 
* It is also apparent that the temperature fluctuates throughout the time measurements, which probably represents the fluctuation of temperature during the day and night cycles. 
* The color of the points seems to go from a dark blue to a lighter blue as time passes, which means that the humidity increases as the measurements go from early to mid January. 
