---
title: "P8105 Homework 1"
author: "Christine Lucille Kuryla"
date: "2023-09-17"
output: github_document
---

The following document is the first homework for P8105.

# Problem 1

First, load `early_january_weather` from the `penguins` dataset in the library `moderndive` as well as the tidyverse.
```{r problem1_load_and_explore, message = FALSE}

# Load moderndive library 
library(moderndive)

# Load tidyverse for dplyr to enable pull when calculating the mean temperature
library(tidyverse)

# This laods the early_january_weather dataset into a variable called early_january_weather
data("early_january_weather")

```

* The variables in this data set were: `r colnames(early_january_weather)`.
  * Some important variables included:
    * `origin`, which was `r pull(early_january_weather,origin)[1]` for every measurement.
    * Year (`year`) and month (`month`), which were the same for every observation: `r pull(early_january_weather,year)[1]` and `r pull(early_january_weather,month)[1]`, respectively.
    * Several weather measurements were included, such as
      * Temperature (`temp`), which ranged from `r min(pull(early_january_weather, temp))` to `r max(pull(early_january_weather, temp))`.
      * Dew point (`dewp`), which ranged from `r min(pull(early_january_weather, dewp))` to `r max(pull(early_january_weather, dewp))`.
      * Humidity (`humid`), which ranged from `r min(pull(early_january_weather, humid))` to `r max(pull(early_january_weather, humid))`.
      * Details about the wind, such as the direction (`wind_dir`), speed (`wind_speed`), and gusts (`wind_gust`).
      * Precipitation (`precip`), which ranged from `r min(pull(early_january_weather, precip))` to `r max(pull(early_january_weather, precip))`.
      * Pressure (`pressure`), which which had some missing values but ranged from `r min(na.omit(pull(early_january_weather, pressure)))` to `r max(na.omit(pull(early_january_weather, pressure)))`.
      * Visibility (`visib`), which had some missing values but ranged from `r min(na.omit(pull(early_january_weather, visib)))` to `r max(na.omit(pull(early_january_weather, visib)))`.
* The data set contained `r nrow(early_january_weather)` observations (rows) with `r ncol(early_january_weather)` variables (columns) describing each observation.
* The mean temperature was `r mean(pull(early_january_weather, temp))`.

```{r problem1_plot}

# Create a scatterplot of temp (y) vs time_hour (x) colored using the humid variable

ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point()

# Save the plot

ggsave("scatter_plot.pdf")

```

In the plot of the temperature vs the date/time, we can see that:

* The data measurements exist from `r min(pull(early_january_weather, time_hour))` to `r max(pull(early_january_weather, time_hour))` and the temperature ranges from `r min(pull(early_january_weather, temp))` to `r max(pull(early_january_weather, temp))`. 
* There is an overall upwards trend in the data, so it seems that as time passes from early to mid January, the temperature warms overall. 
* It is also apparent that the temperature fluctuates throughout the time measurements, which probably represents the fluctuation of temperature during the day and night cycles. 
* The color of the points seems to go from a dark blue to a lighter blue as time passes, which means that the humidity increases as the measurements go from early to mid January. 

# Problem 2

```{r problem2_df_and_means}

# Create a dataframe

dataframe_problem_2 <- tibble(
  samp_norm = rnorm(10),
  greater_than_0 = samp_norm > 0,
  char_vec = c("C", "h", "a", "r", "a", "c", "t", "e", "r", "!"),
  factor_vec = as.factor(c("pink", "green", "blue", "pink", "pink", "pink", "green", "blue", "pink", "green"))
)

# Take the mean of each column of the data frame
mean(pull(dataframe_problem_2, samp_norm))
mean(pull(dataframe_problem_2, greater_than_0))
mean(pull(dataframe_problem_2, char_vec))
mean(pull(dataframe_problem_2, factor_vec))

```

* Taking the mean of the sample created following the normal distribution worked: `r mean(pull(dataframe_problem_2, samp_norm))`
* Taking the mean of the logical vector worked: `r mean(pull(dataframe_problem_2, greater_than_0))`. This was likely the mean of 4 1's (TRUE) and 6 0's (FALSE).
* As expected, taking the mean of the character vector did not work: `r mean(pull(dataframe_problem_2, char_vec))`
* As expected, taking the mean of the factor vector did not work either: `r mean(pull(dataframe_problem_2, factor_vec))`.

```{r problem2_as_numeric, eval = FALSE}

# Apply the as numeric function to the logical (greater_than_0), character (char_vec), and factor vectors (factor_vec)

as.numeric(pull(dataframe_problem_2, greater_than_0))
as.numeric(pull(dataframe_problem_2, char_vec))
as.numeric(pull(dataframe_problem_2, factor_vec))

```

Results of applying the as.numeric function:
* When applied to the logical vector, the as.numeric function converts the vector to 1's for TRUE and 0's for FALSE. 
* For the character vector, the as.numeric function results in all NA, which makes sense because converting characters to numbers is nonsensical. 
* For the factor vector, the as.numeric function results in a vector containing the numbers 1, 2, and 3. This is the rank of the levels of the factors created. 

This explains what happens when we tried to take the mean of the respective vectors. 
* The mean of four 1's and six 0's is 0.4, which corresponds to the mean for the logical vector. 
* The mean of the character vector did not work because that doesn't make sense, and the mean of ten NAs also doesn't make sense. 
* The mean of the factor vector did not work because factors are categorical variables, so a mean doesn't make sense. 
